# Project IV 

##EDA of Red Wine Dataset

*by Xiaopeng Hu*


========================================================




### Quick review of dataset

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load libraries
library(reshape2)
library(plyr)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(knitr)
library(xtable)
library(stringr)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
RedWine_raw<-read.csv('wineQualityReds.csv')

# Review the data
RedWine_raw[!complete.cases(RedWine_raw),]
str(RedWine_raw)

#remove X
RedWine_raw<-RedWine_raw[,2:13]
```


There are 13 variables and 1599 observations in this dataset. The first variable
X is apparently the series of all observations and will not be included in 
following analysis. For other variables, "quality" is the dependent variable, 
so there are 11 independent variables about the physical and chemical properties
of wines, e.g., "fixed.acidity", "volatile.acidity", "citric.acid", 
"residual.sugar", "chlorides",   "free.sulfur.dioxide", "total.sulfur.dioxide",
"density", "pH", "sulphates", "alcohol". The dependent variable "quality" has 
integer attribute, it could also be considered as a categorical variable. All 
independent variables have numeric attribute,

This is a completed dataser without missing value or NA.



# Univariate Plots Section

 

In following analysis, simple statistic analysis are performed on each of the 12
variables, then adjusted binwidths are employed accordingly to produce histogram
plot with more details. 



```{r echo=FALSE,message=FALSE, warning=FALSE }
summary(RedWine_raw$quality)
prop.table(table(RedWine_raw$quality))
# http://stackoverflow.com/questions/6967664/ggplot2-histogram-with-normal-curve
RedWine_raw$quality.factor<-as.factor(RedWine_raw$quality)

ggplot(data = RedWine_raw, 
       aes(x = quality.factor )) +
  geom_histogram(color = "red",
                 fill = "red") +
  xlab("Quality")
 

```


The histogram of "quality" is weakly right-skewed, the median 6.0 is slight 
higher than the mean 5.636. This is of casue a good news for wine producers 
sicne 50% wines have a quality of or aboue 6, and only 4% wines are low 
quality(3 and 4). There are not much high quality wines either, but it is normal
for any products.



```{r echo=FALSE,message=FALSE, warning=FALSE }
summary(RedWine_raw$fixed.acidity) 
summary(RedWine_raw$volatile.acidity) 
summary(RedWine_raw$citric.acid)
summary(RedWine_raw$pH) 

p_fa = ggplot(data = RedWine_raw, 
              aes(x = fixed.acidity)) +
  geom_histogram(aes(y=..density..), 
                 binwidth = 0.1,
                 fill = "red",
                 color = "red")+
  stat_function(fun=dnorm, 
                args=list(mean=mean(RedWine_raw$fixed.acidity),
                                     sd=sd(RedWine_raw$fixed.acidity)))


p_va = ggplot(data = RedWine_raw,
              aes(x = volatile.acidity)) +
  geom_histogram(aes(y=..density..), 
                 binwidth = 0.01, 
                 fill = "red", 
                 color = "red")+
  stat_function(fun=dnorm, 
                args=list(mean=mean(RedWine_raw$volatile.acidity),
                                     sd=sd(RedWine_raw$volatile.acidity)))


p_ca = ggplot(data = RedWine_raw, 
              aes(x = citric.acid)) +
  geom_histogram(aes(y=..density..), 
                 binwidth = 0.01,
                 fill = "red", 
                 color = "red")+
  stat_function(fun=dnorm, 
                args=list(mean=mean(RedWine_raw$citric.acid),
                                     sd=sd(RedWine_raw$citric.acid)))

p_ph = ggplot(data = RedWine_raw, 
              aes(x = pH)) +
  geom_histogram(aes(y=..density..),
                 binwidth = 0.01, 
                 fill = "red",
                 color = "red") +
  stat_function(fun=dnorm, 
                args=list(mean=mean(RedWine_raw$pH),
                                     sd=sd(RedWine_raw$pH)))


grid.arrange(p_fa,p_va,p_ca,p_ph, ncol = 2)

```


"fixed.acidity", "volatile.acidity", "citric.acid" and 'pH' are all about 
acidity of wine, so it is reasonable to analyze them together. "fixed.acidity" 
has left-skewed distribution, while "volatile.acidity" has a rough normal 
distribution. "citric.acid" exhibit a rough uniform distribution with a high 
peak at far left side, which means around 130 tested wines have 0.00 value of 
citric acid. 'pH' has a typical normal distribution.It is interesting to know 
that all these wines have a pH less than 4. Red wines are DRY!



```{r echo=FALSE,message=FALSE, warning=FALSE }
summary(RedWine_raw$residual.sugar)
summary(RedWine_raw$chlorides)

p_rs = ggplot(data = RedWine_raw, 
              aes(x = residual.sugar)) +
  geom_histogram( aes(y=..density..),
                  binwidth = 0.1,
                  fill = "red", 
                  color = "red")+
  stat_function(fun=dnorm, 
                args=list(mean=mean(RedWine_raw$residual.sugar),
                          sd=sd(RedWine_raw$residual.sugar)))

suger_cut_4<-subset(RedWine_raw,RedWine_raw$residual.sugar<4)

p_rs_5 = ggplot(data = suger_cut_4, 
                aes(x = residual.sugar)) +
  geom_histogram( aes(y=..density..),
                  binwidth = 0.1,
                  fill = "red",
                  color = "red")+
  stat_function(fun=dnorm, 
                args=list(mean=mean(suger_cut_4$residual.sugar),
                          sd=sd(suger_cut_4$residual.sugar)))


p_cl = ggplot(data = RedWine_raw, 
              aes(x = chlorides)) +
  geom_histogram(aes(y=..density..),
                 binwidth = 0.001, 
                 fill = "red", 
                 color = "red")+
  stat_function(fun=dnorm, 
                args=list(mean=mean(RedWine_raw$chlorides),
                                     sd=sd(RedWine_raw$chlorides)))


chlorides_cut_2<-subset(RedWine_raw,RedWine_raw$chlorides<0.15)

p_cl_2 = ggplot(data = chlorides_cut_2, 
                aes(x = chlorides)) +
  geom_histogram(aes(y=..density..),
                 binwidth = 0.001, 
                 fill = "red", 
                 color = "red")+
  stat_function(fun=dnorm, 
                args=list(mean=mean(chlorides_cut_2$chlorides),
                          sd=sd(chlorides_cut_2$chlorides)))

grid.arrange(p_rs,p_cl,p_rs_5,p_cl_2,ncol = 2)

```


Both "residual.sugar" and "chlorides" have left-skewed distrubutions and long 
tails at the right side. However, if removes outliers. both exhibit nice normal 
distributions.



```{r echo=FALSE,message=FALSE, warning=FALSE }
summary(RedWine_raw$free.sulfur.dioxide)
summary(RedWine_raw$total.sulfur.dioxide)

p_fsd = ggplot(data = RedWine_raw, 
               aes(x = free.sulfur.dioxide)) +
  geom_histogram(aes(y=..density..),
                 binwidth = 1, 
                 fill = "red",
                 color = "red")+
  stat_function(fun=dnorm, 
                args=list(mean=mean(RedWine_raw$free.sulfur.dioxide),
                          sd=sd(RedWine_raw$free.sulfur.dioxide)))

p_tsd = ggplot(data = RedWine_raw, 
               aes(x = total.sulfur.dioxide)) +
  geom_histogram(aes(y=..density..),
                 binwidth = 3, 
                 fill = "red", 
                 color = "red")+
  stat_function(fun=dnorm, 
                args=list(mean=mean(RedWine_raw$total.sulfur.dioxide),
                          sd=sd(RedWine_raw$total.sulfur.dioxide)))

grid.arrange(p_fsd,p_tsd,ncol = 2)

```


Both "free.sulfur.dioxide" and "total.sulfur.dioxide" have left-skewed distributions.




```{r echo=FALSE,message=FALSE, warning=FALSE}
summary(RedWine_raw$density)
ggplot(data = RedWine_raw, 
       aes(x = density)) +
  geom_histogram(aes(y=..density..),
                 binwidth = 0.0001, 
                 fill = "red", 
                 color = "red")+
  stat_function(fun=dnorm, 
                args=list(mean=mean(RedWine_raw$density),
                          sd=sd(RedWine_raw$density)))

```


The distribution for "density has a normal appearence. 




```{r echo=FALSE,message=FALSE, warning=FALSE}
summary(RedWine_raw$sulphates)
ggplot(data = RedWine_raw, 
       aes(x = sulphates)) +
  geom_histogram(aes(y=..density..),
                 binwidth = 0.02, 
                 fill = "red", 
                 color = "red")+
  stat_function(fun=dnorm, 
                args=list(mean=mean(RedWine_raw$sulphates),
                          sd=sd(RedWine_raw$sulphates)))

```


The distribution for "sulphates" has a normal appearence with outliers at the 
right side.


  
```{r echo=FALSE,message=FALSE, warning=FALSE}
summary(RedWine_raw$alcohol)
ggplot(data = RedWine_raw, 
       aes(x = alcohol)) +
  geom_histogram(aes(y=..density..),
                 binwidth = 0.1, 
                 fill = "red", 
                 color = "red") +
  stat_function(fun=dnorm, 
                args=list(mean=mean(RedWine_raw$alcohol),
                          sd=sd(RedWine_raw$alcohol)))

```


The distribution for "alcohol" is left skewed.



# Univariate Analysis



### What is the structure of your dataset?

This completed dataset contains 13 variables and 1599 observations. Variable "X"
is indicator of observations. Integer variable "quality" is the median of wine 
qualities based on at least 3 wine experts evaluations; the rest 11 numeric 
variables are about the physical and chemical properties of wines. 


### What is/are the main feature(s) of interest in your dataset?

Correlationship between quality and the 11 physical and chemical properties of 
wines is the main feature of interest in the dataset. However, quality is scored
by personals, humans can not be as accurate as equilpments, and the final 
quality score is a median of more than three scores, so each quality level is 
very likely to cover a wide range of each physical and chemical properties of 
wine. The correlationship may be complicated. 


### What other features in the dataset do you think will help support your 
### investigation into your feature(s) of interest?

Some independent variables are actually related with eacher other, such as "pH" 
and "fixed.acidity"", "volatile.acidity"", "citric.acid". These interior 
correlations could be used to cross-verify dataset. 

### Did you create any new variables from existing variables in the dataset?

No.

### Of the features you investigated, were there any unusual distributions? Did 
#### you perform any operations on the data to tidy, adjust, or change the form 
#### of the data? If so, why did you do this?

"citric.acid" exhibits a rough uniform distribution, while others exhibit normal
or skewed distributions. No operations was performed to tidy the dataset at this
early stage.




# Bivariate Plots Section

For quick reviewing the correlationships between all 12 variables, a correlation
table is produced with cor() function (ggplairs is tedius for this purpose). 


```{r echo=FALSE, message=FALSE, warning=FALSE}
 
corr<-cor(x=RedWine_raw[,1:12], y=RedWine_raw[,1:12])

corr<-as.matrix(corr)

corr[lower.tri(corr,diag = TRUE)] <- NA

kable(corr,  format = "markdown")

corr.m<-melt(corr)

ggplot(corr.m,
       aes(x = Var1, 
           y = Var2,
           fill = value)) + 
  geom_tile() + 
  scale_fill_gradient(low = "blue",
                      high = "red",
                      na.value="white") + 
  xlab("") + 
  ylab("") +
  theme(axis.text.x = element_text(angle=60, hjust=1))
```


As expected, the 11 independent variables exhibit none to moderate correlations
with the dependent variable "quality". The variable having the highest 
correlation of 0.476 with "quality" is "alcohol", the second is 
"volatile.acidity".  It is also confirmed that each quality score covers a wide 
range of each single physical and chemical properties of wine.



```{r echo=FALSE, message=FALSE, warning=FALSE}

p_fa_q = ggplot(data = RedWine_raw, 
                aes(x = volatile.acidity, 
                    y = quality)) + 
geom_jitter(alpha = 0.1,
              position = position_jitter(height = .2)) +
  geom_smooth( method = 'lm')  
  

p_al_q = ggplot(data = RedWine_raw, 
                aes(x = alcohol, 
                    y = quality)) + 
  geom_jitter(alpha = 0.1,
              position = position_jitter(height = .2)) +
  geom_smooth( method = 'lm')

grid.arrange(p_fa_q,p_al_q,ncol = 2)

```


For a clear view, correlation plots of  "alcohol" and "volatile.acidity" with 
"quality" are produced here.



```{r echo=FALSE, message=FALSE, warning=FALSE}

p_fa_p = ggplot(data = RedWine_raw, 
                aes(x = fixed.acidity, 
                    y = pH)) +
  geom_point(alpha = 0.1) +
  geom_smooth( method = 'lm')+
  scale_x_log10()

p_ca_fc = ggplot(data = RedWine_raw, 
                 aes(x = citric.acid, 
                     y = fixed.acidity)) +
  geom_point(alpha = 0.1 ) +
  geom_smooth( method = 'lm')

p_fsd_tsd = ggplot(data = RedWine_raw, 
                   aes(x = free.sulfur.dioxide, 
                       y = total.sulfur.dioxide)) +
  geom_point(alpha = 0.1) +
  geom_smooth( method = 'lm')

p_fa_d = ggplot(data = RedWine_raw, 
                aes(x = fixed.acidity, y = density)) +
  geom_point(alpha = 0.1) +
  geom_smooth( method = 'lm')


grid.arrange(p_fa_p,p_ca_fc,p_fsd_tsd,p_fa_d,ncol = 2)

```


On the other hand, interior correlations do exist. 

"pH" and "fixed.acidity" have a strong negtive correlation of -0.668, which 
makes sense since pH is roughly defined as the negative of the logarithm to 
base 10 of the concentration of hydrogen ions released by acids in solution. 

"fixed.acidity" and "citric.acid" have a strong positive correlation of 0.672, 
which also makes sense since ciitric acid is one the major acids in fruits. 
Indeed, "pH" does have a correlation of -0.542 with "citric.acid".  

"total.sulfur.dioxide" and "free.sulfur.dioxide" have a strong positive 
correlation of 0.668, because free form sulfur dioxide exists in equilibrium 
between dissolved sulfur dioxide and sulfite ion, e.g., the total sulfur dioxide. 

"density" and "fixed.acidity" also have a strong positive correlation of 0.668,
which  makes sense because organic acids such as citric acid have higher density
than water.






Because "quality" can be considered as categery variable, it is interesting to 
group the 11 independent variables according to "quality" level,  then to check 
the relationships of means/medians of levels and coresponding quality levels. 


```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x=quality,y=alcohol),
       data=RedWine_raw) +
    geom_jitter(alpha = 0.3,
              color = "black",
              position = position_jitter(width = .2)) +
  geom_line(stat = "summary",
            fun.y = mean,
            color = "black") +
  geom_line(stat = "summary",
            fun.y = median,
            color = "green") +
  geom_line(stat = "summary",
            fun.y = quantile, 
            probs = 0.1,
            linetype = 2,
            color = 'blue') +
  geom_line(stat = "summary",
            fun.y = quantile, 
            probs = 0.9,
            linetype = 2,
            color = 'blue')


```


Very interestingly, the means/medians of "alcohol" of "quality" of levels 
exhibit strong linear correlation with "quality" levels. I therefore suspect 
that this observation is also true for other independent variables. 



```{r echo=FALSE, message=FALSE, warning=FALSE}
RW.mean_by_quality<- RedWine_raw %>%
  group_by(quality) %>%
  summarise_each(funs(mean))

```


To prove this assuming, a new dataframe containing means of all 11 indenpendent 
variables grouped by "quality" levels is produced. A correlation table of 
"quality" v.s means of 11 indenpendent variables grouped by "quality" is 
generated accodingly and is compared with that of "quality" v.s. 11 indenpendent
variables. 


```{r echo=FALSE, message=FALSE, warning=FALSE }
 
corr.mean<-cor(x=RW.mean_by_quality[,2:12], y=RW.mean_by_quality$quality)
corr.mean<-t(corr.mean)
corr.mean<-as.data.frame(corr.mean)
 
corr.mean.m<-melt(corr.mean)
 


corr.qua<-cor(x=RedWine_raw[,1:11], y=RedWine_raw$quality)

corr.qua<-t(corr.qua)
corr.qua<-as.data.frame(corr.qua)
corr.qua.m<-melt(corr.qua)

corr.comp<-merge(corr.mean.m,corr.qua.m, by = "variable")

colnames(corr.comp)<-c("property", "means.by.group", "raw")

corr.comp<-corr.comp[order(-abs(corr.comp$means.by.group)),]

kable(corr.comp,  format = "markdown")

corr.comp.m<- melt(corr.comp)

ggplot(corr.comp.m, aes(property, value, fill=variable))  + 
  geom_bar(position="dodge",stat="identity") +
  ylab("correlation coefficient") +
  theme(axis.text.x = element_text(angle=60, hjust=1))
 
```


I can see that by grouping with "quality", almost the means of indenpendent 
variables exhibit strong linear correlation with "quality" levels. The following
analysis will concentrate in the top six variables, e.g., "alcohol", "sulphates"
,"pH", "chlorides", "volatile.acidity" and "citric.acid", with  correlation 
coefficient greater than 0.9.



```{r echo=FALSE, message=FALSE, warning=FALSE}

c_s = ggplot(data = RW.mean_by_quality, aes(x =sulphates, 
                                            y = quality )) +
  geom_point() +
  geom_smooth(method = 'lm')

c_c = ggplot(data = RW.mean_by_quality, aes(x =citric.acid, 
                                            y = quality )) +
  geom_point() +
  geom_smooth(method = 'lm')

c_v = ggplot(data = RW.mean_by_quality, aes(x =volatile.acidity, 
                                            y = quality )) +
  geom_point() +
  geom_smooth(method = 'lm')

c_p = ggplot(data = RW.mean_by_quality, aes(x =pH, 
                                            y = quality )) +
  geom_point() +
  geom_smooth(method = 'lm')

c_ch = ggplot(data = RW.mean_by_quality, aes(x =chlorides, 
                                             y = quality )) +
  geom_point() +
  geom_smooth(method = 'lm')

c_a = ggplot(data = RW.mean_by_quality, aes(x =alcohol, 
                                            y = quality )) +
  geom_point() +
  geom_smooth(method = 'lm')

grid.arrange(c_s, c_c, c_p, c_v, c_ch, c_a, ncol = 2)

```


A similiar conclusion can be obtained with medians. 



# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the 
### investigation. How did the feature(s) of interest vary with other features  
### in the dataset?  

"aclcohol" has modedate correlation with "quality", others have weak or none 
correlations with "quality". However, The means/medians of "aclcohol", 
"sulphates" and "citric.acid" of each "quality" levels exhibit very strong 
positive linear correlation with "quality", while means/medians of of "pH", 
"chlorides" and "volatile.acidity" exhibit very strong negative linear 
correlation with "quality".

### Did you observe any interesting relationships between the other features 
### (not the main feature(s) of interest)?

There are strong interior correlations between "pH", "volatile.acidity" and 
"citric.acid", "total.sulfur.dioxide" and "free.sulfur.dioxide" aslo have strong
interior correlations. In regression analysis, these interior correlations need 
to be counted.What was the strongest relationship you found


### What was the strongest relationship you found?

"aclcohol" and "quality" have  the strongest correlation between independent and
dependent variables, while "pH" and "fixed.acidity" have  the strongest interior
correlation among independent variables.


# Multivariate Plots Section

As seen in the analysis above, in all 11 indenpendent variables,"alcohol" has 
the highest but moderate correlation coefficient of 0.476 to wine quality, while
after being grouped by "quality", the means/medians of "sulphates","pH", 
"chlorides", "volatile.acidity", "citric.acid" and "alcohol" exhibit strong 
correlation with wine quality. Therefore I produced bivariate scatterplots of 
"alcohol" with other indenpendent vraibales to see if we can find stronger
relationships between two indenpendent vraibales and the dependent vraibale 
"quality",   



```{r fig.width = 12, fig.height = 10, echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}


b.a.s = ggplot(data = RedWine_raw, 
               aes(x = alcohol, 
                   y = sulphates, 
                   color = quality.factor)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =1) +
  scale_color_brewer(type = "seq") 

b.a.va = ggplot(data = RedWine_raw, 
                aes(x = alcohol, 
                    y = volatile.acidity,
                    color = quality.factor)) +
  geom_point(alpha = 0.3)  +
  stat_smooth(method="lm",
              size =1) +
  scale_color_brewer(type = "seq")

b.a.cl = ggplot(data = RedWine_raw, 
                aes(x = alcohol, 
                    y = chlorides,
                    color = quality.factor)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =1) +
  scale_color_brewer(type = "seq")

b.a.ca = ggplot(data = RedWine_raw, 
                aes(x = alcohol, 
                    y = citric.acid, 
                    color = quality.factor)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =1) +
  scale_color_brewer(type = "seq")

grid.arrange(b.a.s, b.a.va, b.a.cl, b.a.ca, ncol = 2)

```


It is clear that spots of different qualities don't show clear ranges. However, 
spots with quality under 5 and these above 5 do exhibit some aggregation. This 
observation implies that the dependent variable "quality" is not sensitive to 
the independent variables. 



```{r fig.width = 12, fig.height = 10,echo=FALSE, message=FALSE, warning=FALSE}

RedWine_raw$rating <- ifelse(RedWine_raw$quality < 5, 
                             'low', 
                             ifelse(RedWine_raw$quality < 7, 
                                    'average', 'high'))


RedWine_raw.lh<-subset(RedWine_raw, RedWine_raw$rating != "average" )

b.a.s.lh = ggplot(data = RedWine_raw.lh, 
                  aes(x = alcohol, 
                      y = sulphates,
                      color = rating)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =1)

b.a.va.lh = ggplot(data = RedWine_raw.lh, 
                   aes(x = alcohol,
                       y = volatile.acidity,
                       color = rating)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =1)

b.a.cl.lh = ggplot(data = RedWine_raw.lh, 
                   aes(x = alcohol, 
                       y = chlorides,
                       color = rating)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =1)

b.a.ca.lh = ggplot(data = RedWine_raw.lh, 
                   aes(x = alcohol, 
                       y = citric.acid, 
                       color = rating)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =1)

grid.arrange(b.a.s.lh, b.a.va.lh, b.a.cl.lh, b.a.ca.lh, ncol = 2)

```


If wines with quality less than 4 are defined as "low quality", greater than 6 
as "high quality", any in the middle as "average quality", bivariate 
scatterplots of "alcohol" with "sulphates", "volatile.acidity" and "citric.acid"
can distinguish "low quality" from "high quality", but not "average quality". 



```{r fig.width = 12, fig.height = 10, echo=FALSE, message=FALSE, warning=FALSE }

b.a.fa.lh = ggplot(data = RedWine_raw.lh, 
                   aes(x = alcohol, 
                       y = fixed.acidity, 
                       color = rating)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =1)

b.a.rs.lh = ggplot(data = RedWine_raw.lh, 
                   aes(x = alcohol, 
                       y = residual.sugar,
                       color = rating)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =1)

b.a.fsd.lh = ggplot(data = RedWine_raw.lh, 
                    aes(x = alcohol, 
                        y = free.sulfur.dioxide,
                        color = rating)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =1)

b.a.tsd.lh = ggplot(data = RedWine_raw.lh, 
                    aes(x = alcohol,
                        y = total.sulfur.dioxide,
                        color = rating)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =1)

grid.arrange(b.a.fa.lh, b.a.rs.lh, b.a.fsd.lh, b.a.tsd.lh, ncol = 2)

```


Bivariate scatterplots of "alcohol" with independent variables with week/none 
correlation coefficentss generally have weaker power to distinguish  "low quality" from 
"high quality".



```{r fig.width = 12,echo=FALSE, message=FALSE, warning=FALSE }

b.rs.tsd = ggplot(data = RedWine_raw.lh, 
                  aes(x =  residual.sugar, 
                      y = total.sulfur.dioxide, 
                      color = rating)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =1) +
  xlim(0,8)
 
b.rs.fsd = ggplot(data = RedWine_raw.lh, 
                  aes(x =  residual.sugar, 
                      y = free.sulfur.dioxide,
                      color = rating)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =1) +
  xlim(0,8)
  

grid.arrange(b.rs.tsd,b.rs.fsd, ncol = 2)

```


Bivariate scatterplots of independent variables with week/none correlation 
coefficents can't distinguish "low quality" wines from "high quality" wines.


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the 
### investigation. Were there features that strengthened each other in terms of 
### looking at your feature(s) of interest?

Among all 11 independedt variables, "alcohol" plays a dominant but still weak 
role in correlations with the dependent variable of "quality", while independedt
variables of "sulphates", "volatile.acidity" and "citric.acid" indeed strengthen
the correlations.


### Were there any interesting or surprising interactions between features?

It is interesting to see that bivariate scatterplots with certain independent 
variables can distingusidh low/high end values of dependent variable. I guess it
is either caused by wide mesurement error of wine quality, or there are other 
important wine pproperties such as fragrance are not included in this dataset.

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths
### and limitations of your model.

No, I didn't. The correlations between indenpendent variables and dependent 
variable are too weak, model based on current analysis is very likey meaningless
. However, complex model might exist.

------

# Final Plots and Summary

### Plot One

```{r  echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
ggplot(corr.m,aes(x = Var1, 
                  y = Var2,
                  fill = value))+ 
  geom_tile() + 
  scale_fill_gradient(low = "blue",
                      high = "red",
                      na.value="white") + 
  xlab("") + 
  ylab("") +
  theme(axis.text.x = element_text(angle=60, hjust=1)) +
  ggtitle("Correlations of the 11 Independent Variables with Wine Quality")

```



### Description One

One of the major purpose of performing EDA with dataset is to investigate the correlations among independent and dependent variables and provide a general map
for advanced analysis such as regression modeling. Plot one demonstrates 
correlations of 11 indenpendent variables with wine quality. With Plot one, I 
can find that content of alcohol has a moderate but the highest correlation coeffeicient. volatile acidity, sulphates and citric acid follow with it. Other independent variables have weak or none correlations with wine quality. There 
are also strong interior correlations between "pH"/"fixed.acidity", and "citric.acid"/"fixed.acidity", "density"/"fixed.acidity", and "total.sulfur.dioxide"/"free.sulfur.dioxide". Therefore, my analysis can focus 
on the correlations between "quality" and "alcohol".



### Plot Two

```{r fig.width = 10, echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}

ggplot(aes(x=quality,
           y=alcohol),
       data=RedWine_raw) +
  geom_jitter(alpha = 0.3,
              color = "black",
              position = position_jitter(width = .2)) +
  geom_line(aes(color = "Mean"),
            stat = "summary",
            fun.y = mean) +
  geom_line(aes(color = "Median"),
            stat = "summary",
            fun.y = median) +
  geom_line(aes(color = "10% quantile"),
            stat = "summary",
            fun.y = quantile, 
            probs = 0.1,
            linetype = 2) +
  geom_line(aes(color = "90% quantile"),
            stat = "summary",
            fun.y = quantile, 
            probs = 0.9,
            linetype = 2) +
  geom_smooth(data = RW.mean_by_quality, 
              aes(y =alcohol, 
                  x = quality, color = "Liner regression of means/levels" ), 
              method = 'lm', size =1) +
  ylab("Alcohol %") +
  xlab("Quality level") +
  ggtitle("Strong Dependence of QUALITY Levels with Means of ALCOHOL Grouped by QUALITY") + 
  theme(legend.title=element_blank())

```


### Description Two

In bivariate analysis, simple scatter spots plots didn't result much useful information about correlations between independent/dependent variables. The main reson could be the defination of wine quality, which may cover a range of indenpendent varibales distribution. Plot Two indeed reveales that if "alcohol" is grouped by "quality", then 
means/medians of each sungroups have strong denpendence with "quality".   

### Plot Three

```{r  fig.width = 10,echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}

ggplot(data = RedWine_raw.lh, aes(x = alcohol,
                                  y = citric.acid, 
                                  color = rating))  +
  geom_point(alpha = 0.3) +
  stat_smooth(method="lm",
              size =2) +
  xlab("Alcohol %") +
  ylab("Citric acid g/L") +
  ggtitle(str_wrap("Binding of Alcohol and Citric acid Can Distinguish
          Low/High Quality of Wine"))




```


### Description Three

Since each qualitu level covers a wide range of independent variables 
distrubution, it actually means the depedent variable "qualitly" is not 
sensitive to independent variables. Plot Three demonstrates that combination of 
two independent variables, "citric.acid" and "alcohol", which have relative high 
correlation coefficents with "quality", can roughly distinguish wines with low
quallity (with "quality" <5) and wines with high quallity (with "quality" >7).

------

# Reflection

Through this exploratory data analysis, I identified key factors related with 
wine quality. Among them, alcohol content plays a major role, while acidity and
sulphates also contribute. However, there must be many other factors that are 
related with wine quality. Altough graphs can illustrate relations between these
factors and wine quality, more sophisticated statistic analysis are definetly 
needed to get a more consolid conslusions. On the other hand, wine quality is 
decided by many independent factors, it is apparently impossible to distinguish
wine quality just by combinding of two factors. Combing of more independent 
variables is likely to give a better results, but this kind of analysis needs to
be done in high dimesional space and is really beyond the scope of current 
project.

The hard part of this EDA comes from the poor correlations between 
independent/dependent variables. Although I have done data analysis for a long 
time in chmistry/drug areas, this is the first time I saw such a dataset. Really
really confusing! Fortunately with google, I reallized that the wine quality is 
not like any other "quality" I met before and it is better to treat wine quality
as level or categary variable rather than numeric variable. Actually If I 
carefully read the instruction I should find out that "quality" is the average
of at least three evaluation results, this is an important clue but I just didnt
pay enough attentions. Yes, dog sometimes must open eyes, nose may be misleading!
